---
title: "p8131_hw4_jsg2145"
author: "Jared Garfinkel"
date: "2/20/2020"
output: pdf_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(nnet)
library(MASS)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

```{r}
df_low = tibble(
  conx = rep(c("low", "high"), 3),
  hous_tp = rep(c("tr_bk", "apt", "hous"), each = 2),
  values = c(65, 34, 130, 141, 67, 130)
) %>% 
  mutate(sat = "low") 

df_med = tibble(
  conx = rep(c("low", "high"), 3),
  hous_tp = rep(c("tr_bk", "apt", "hous"), each = 2),
  values = c(54, 47, 76, 116, 48, 105)
) %>% 
  mutate(sat = "med")

df_high = tibble(
  conx = rep(c("low", "high"), 3),
  hous_tp = rep(c("tr_bk", "apt", "hous"), each = 2),
  values = c(100, 100, 111, 191, 62, 104)
) %>% 
  mutate(sat = "high")

df_sum = union(df_low, df_med) %>% 
  union(df_high) %>% 
  unnest() %>% 
  pivot_wider(names_from = sat, values_from = values)

df_tbl = df_sum %>% 
  group_by(conx, hous_tp) %>% 
  mutate(n = sum(low, med, high)) %>% 
  summarize(low = low/n,
            med = med/n,
            high = high/n)
```

```{r}
df_conx = df_sum %>%
  group_by(conx) %>% 
  summarize(low = sum(low),
            med = sum(med),
            high = sum(high),
            n = sum(low, med, high)) %>% 
  group_by(conx) %>% 
  summarize(low = low/n,
            med = med/n,
            high = high/n)
```

```{r}
df_hous_tp = df_sum %>%
  group_by(hous_tp) %>% 
  summarize(low = sum(low),
            med = sum(med),
            high = sum(high),
            n = sum(low, med, high)) %>% 
  group_by(hous_tp) %>% 
  summarize(low = low/n,
            med = med/n,
            high = high/n)
```


```{r}
df_tidy_tbl = df_tbl %>% 
  pivot_longer(cols = c(low, med, high), names_to = "sat", values_to = "pct") %>% 
  ungroup() %>% 
  mutate(sat = str_remove(sat, "_pct$"),
         sat = factor(sat, labels = c("low", "med", "high")),
         conx = factor(conx, labels = c("low", "high")),
         hous_tp = factor(hous_tp, labels = c("apt", "hous", "tr_bk")))
```


```{r, results = "hide"}
df_tidy_tbl %>% 
  ggplot(aes(x = sat, y = pct)) +
  geom_col() +
  facet_grid(hous_tp~conx)
```

```{r, results = "hide"}
df_conx %>% 
  pivot_longer(cols = low:high, names_to = "sat") %>% 
  ggplot(aes(x = conx, y = value, fill = sat)) +
  geom_bar(position = "dodge", stat = "identity")
```

```{r, results = "hide"}
df_hous_tp %>% 
  pivot_longer(cols = low:high, names_to = "sat") %>% 
  ggplot(aes(x = hous_tp, y = value, fill = sat)) +
  geom_bar(position = "dodge", stat = "identity")
```


```{r}
hagen.multi = nnet::multinom(cbind(low, med, high) ~ conx + hous_tp, data = df_sum)

summary(hagen.multi)
```

It appears that low contact with neighbors is associated with lower medium and high satisfaction holding housing type constant, while higher contact with neighbors is associated with an increase in medium and high satisfaction in the tower block housing compared to apartments. But, there is no clear trend between satisfaction and house type between houses and apartments while holding contact constant. 

```{r}
pihat = predict(hagen.multi, type = "probs") 
m = rowSums(df_sum[3:5])
res.pearson = (df_sum[3:5] - pihat * m) / sqrt(pihat * m)
```

```{r}
G.stat = sum(res.pearson ^ 2) # Generalized Pearson Chisq Stat
G.stat
pval = 1 - pchisq(G.stat, df = (6 - 4) * (3 - 1))
pval
```


The p-value is `r round(pval, digits =3)`, so we do not reject the null, indicating the fit is acceptable.

```{r tidy mod}
tidy.err = summary(hagen.multi)$standard.error %>% 
  broom::tidy() %>% 
  mutate(value = "err")

tidy.coef = summary(hagen.multi)$coefficients %>% 
  broom::tidy() %>% 
  mutate(value = "coef")

tidy.multi = union(tidy.coef, tidy.err) %>% 
  janitor::clean_names() %>% 
  dplyr::select(-x_intercept) %>% 
  pivot_longer(cols = c(conxlow, hous_tphous, hous_tptr_bk), values_to = "estimate") %>% 
  pivot_wider(names_from = value, values_from = estimate)
```


```{r}
tidy.ci = tidy.multi %>% 
  group_by(rownames, name) %>% 
  mutate(
    lower = (coef - qnorm(.975) * err),
    upper = (coef + qnorm(.975) * err)) %>%
  dplyr::select(rownames, name, lower, upper) %>% 
  ungroup() %>% 
  rename(sat = rownames, par = name)

tidy.ci %>% 
  knitr::kable(digits = 3)
```

The odds ratio of reporting medium satisfaction compared to low satisfaction is `r filter(tidy.multi, rownames == "med", name == "conxlow") %>% pull(coef) %>% round(digits = 3) %>% exp()`, CI95%: (`r filter(tidy.ci, sat == "med", par == "conxlow") %>% pull(lower) %>% round(digits = 3) %>% exp()`, `r filter(tidy.ci, sat == "med", par == "conxlow") %>% pull(upper) %>% round(digits = 3) %>% exp()`) for low contact compared to high contact holding house type constant.

The odds ratio of reporting medium satisfaction compared to low satisfaction is `r filter(tidy.multi, rownames == "med", name == "hous_tphous") %>% pull(coef) %>% round(digits = 3) %>% exp()` CI95%: (`r filter(tidy.ci, sat == "med", par == "hous_tphous") %>% pull(lower) %>% round(digits = 3) %>% exp()`, `r filter(tidy.ci, sat == "med", par == "hous_tphous") %>% pull(upper) %>% round(digits = 3) %>% exp()`) for people living in houses compared to those living in apartments holding contact constant.

The odds ratio of reporting medium satisfaction compared to low satisfaction is `r filter(tidy.multi, rownames == "med", name == "hous_tptr_bk") %>% pull(coef) %>% round(digits = 3) %>% exp()` CI95%: (`r filter(tidy.ci, sat == "med", par == "hous_tptr_bk") %>% pull(lower) %>% round(digits = 3) %>% exp()`, `r filter(tidy.ci, sat == "med", par == "hous_tptr_bk") %>% pull(upper) %>% round(digits = 3) %>% exp()`) for people living in tower blocks compared to those living in apartments holding contact constant.

The odds ratio of reporting high satisfaction compared to low satisfaction is `r filter(tidy.multi, rownames == "high", name == "conxlow") %>% pull(coef) %>% round(digits = 3) %>% exp()` CI95%: (`r filter(tidy.ci, sat == "high", par == "conxlow") %>% pull(lower) %>% round(digits = 3) %>% exp()`, `r filter(tidy.ci, sat == "high", par == "conxlow") %>% pull(upper) %>% round(digits = 3) %>% exp()`) for people who reported low contact compared to those who reported high contact holding housing type constant.

The odds ratio of reporting high satisfaction compared to low satisfaction is `r filter(tidy.multi, rownames == "high", name == "hous_tphous") %>% pull(coef) %>% round(digits = 3) %>% exp()` CI95%: (`r filter(tidy.ci, sat == "high", par == "hous_tphous") %>% pull(lower) %>% round(digits = 3) %>% exp()`, `r filter(tidy.ci, sat == "high", par == "hous_tphous") %>% pull(upper) %>% round(digits = 3) %>% exp()`) for people who live in houses compared to those who live in apartments holding contact constant.

The odds ratio of reporting high satisfaction compared to low satisfaction is `r filter(tidy.multi, rownames == "high", name == "hous_tptr_bk") %>% pull(coef) %>% round(digits = 3) %>% exp()` CI95%: (`r filter(tidy.ci, sat == "high", par == "hous_tptr_bk") %>% pull(lower) %>% round(digits = 3) %>% exp()`, `r filter(tidy.ci, sat == "high", par == "hous_tptr_bk") %>% pull(upper) %>% round(digits = 3) %>% exp()`) for people who live in tower blocks compared to those who live in apartments holding contact constant.

## Part iii

```{r}
# proportional odds model
freq = df_sum %>%
  pivot_longer(cols = c(low, med, high), names_to = "sat")

df_ord = df_tidy_tbl %>% 
  left_join(freq, by = c("sat", "conx", "hous_tp")) %>% 
  mutate(sat = factor(sat, 
                      levels = c("low", "med", "high"), 
                      ordered = TRUE))

# fit proportional odds model
df_polr = polr(sat ~ conx + hous_tp, data = df_ord, weights = value)

summary(df_polr)
```

The change in log odds of falling into a lower satisfaction category associated with low contact is `r coef(df_polr) %>% broom::tidy() %>% filter(names == "conxlow") %>% dplyr::select(x) %>% as.numeric() %>% round(digits = 3)` compared to high contact holding housing type constant.

The change in log odds of falling into a lower satisfaction category associated with living in a house is `r coef(df_polr) %>% broom::tidy() %>% filter(names == "hous_tphous") %>% dplyr::select(x) %>% as.numeric() %>% round(digits = 3)` compared to those living in apartments holding level of contact constant.

The change in log odds of falling into a lower satisfaction category associated with living in a tower block is `r coef(df_polr) %>% broom::tidy() %>% filter(names == "hous_tptr_bk") %>% dplyr::select(x) %>% as.numeric() %>% round(digits = 3)` compared to those living in apartments holding level of contact constant.

## Part iv

```{r}
# Pearson residuals for proportional odds model

pihat = predict(df_polr, df_sum, type = 'p')
m = df_sum %>% 
  group_by(conx, hous_tp) %>% 
  mutate(total = sum(low, med, high)) %>% 
  ungroup() %>% 
  dplyr::select(total) %>% 
  simplify()
res.pearson = (df_sum[,3:5] - pihat * m) / sqrt(pihat * m)
G = sum(res.pearson^2)
G
numsamp = (3 - 1) * 6 # degree of freedom for grouped data
numparam = 2 + 3 # total num of param
pval = 1 - pchisq(G, df = numsamp - numparam)
pval
```
The p-value is `r round(pval, digits = 3)` (>0.05), indicating that the model is a good fit.

```{r}
## find the discrepancies

bind_cols(df_sum, res.pearson) %>% 
  dplyr::select(conx, hous_tp, low1:high1)
```


It appears that much of the discrepancy between the observed and expected models are in the low contact group and the group living in houses. However, there is also a high residual in the group with high contact living in the tower blocks.